Casual Bandits

### 2016-01-13 causal-bandit/icml-2015/simple-regret.tex
incorporated the simple-regret upper bound and continued to cleanup the proof

\begin{theorem}\label{thm:unknown_q_simpleregret}
Define $m = \min\set{1 \leq i \leq N:q_{i+1} \geq \frac{1}{i}}$.
Then Algorithm \ref{alg:simple} satisfies
\eq{
r_T \in \bigo{\sqrt{\frac{m}{T}\log\left(\frac{NT}{m}\right)}}\,.
}
\end{theorem}

@@

\todot{Do we have a proof for this in another doc?}
\begin{proof}
\end{proof}

>>>>

\begin{definition}
Let $\boldsymbol{q} \in [0,1]^N$. Then define $q^{(i)}$ to be the $i$th largest value of the set $\set{\min \set{q_j, 1 - q_j} : 1 \leq j \leq N}$.
For example, if $\boldsymbol{q} = (0,\, 2/3,\, 1/2)$, then $q^{(1)} = 0$, $q^{(2)} = 1/3$, $q^{(3)} = 1/2$.
Then define 
\eqn{
\label{def:m}
m(\boldsymbol{q}) = \min\set{1 \leq i \leq N: q^{(i+1)} \geq \frac{1}{i}}\,.
}
\end{definition}

\begin{theorem}\label{thm:unknown_q_simpleregret}
Algorithm \ref{alg:simple} satisfies
\eq{
\simpleregret \in \bigo{\sqrt{\frac{m}{T}\log\left(\frac{NT}{m}\right)}}\,.
}
\end{theorem}

@@

The proof of Theorem \ref{thm:uq-simple} requires some lemmas, the first of which is an immediate consequence of the Chernoff bound.

\begin{lemma}\label{lem:conc1}
Let $i \in \set{1,\ldots, N}$ and $\delta > 0$. Then
\eq{
\P{\left|\hat q_i - q_i\right| \geq \sqrt{\frac{6q_i}{T} \log \frac{2}{\delta}}} \leq \delta\,.
}
\end{lemma}

\iffalse
\begin{proof}
Let $Z_t = \ind{X_{i,t} = 1} \in \set{0,1}$.
Then
\eq{
\hat q_i = \frac{2}{T} \sum_{t=1}^{T/2} Z_t\,.
}
Now $Z_1,\ldots,Z_{T/2}$ is an i.i.d.\ sequence of Bernoulli random variables with with mean $q_i$. The result follows from the Chernoff bound.
\end{proof}
\fi

\begin{lemma}\label{lem:m_est}
Let $\delta >0$ and assume $T \geq 48m \log\frac{4N}{\delta}$. Then
\eq{
\P{\hat m < \frac{2}{3}m} \leq \delta\ \text{ and } \P{\hat m > 2m} \leq \delta\,.
}
\end{lemma}

\begin{proof}
Let $\boldsymbol{q}^b$ and $\boldsymbol{q}^{ub}$ and be the maximally balanced and unbalanced $\boldsymbol{q}$ for a given $m$ respectively.
\eq{
q^{ub}_i = \begin{cases}
0 & \text{if } i \leq m \\
\frac{1}{m} & \text{otherwise}\,.
\end{cases} \\
q^b_i < \begin{cases}
\frac{1}{m} & \text{if } i \leq m \\
1 & \text{otherwise}\,.
\end{cases}  
}
For $\hat{m}$ to over-estimate $m$, we must identify some balanced arms as unbalanced. For $\hat{m}$ to under-estimate $m$, we must identify some unbalanced arms as balanced.
\eq{
\P{\hat m > 2m} \leq \P{\hat m > 2m|\boldsymbol{q} = \boldsymbol{q}^{ub}}\\
\P{\hat m < \frac{2}{3}m} \leq \P{\hat m < \frac{2}{3}m|\boldsymbol{q} = \boldsymbol{q}^{b}}
}

Given $\boldsymbol{q} = \boldsymbol{q}^{ub}$, we have by Lemma \ref{lem:conc1}, with probability at least $1 - \delta$ that:

\eq{
& \left| \hat q_i - q_i\right| 
 \leq \begin{cases}
0 & \text{if } i \leq m\\
\sqrt{\frac{6}{mh} \log\frac{2}{\delta}} & \text{otherwise} \\
\end{cases}\\
\implies &
\begin{cases}
(\forall i \leq m) \qquad \left| \hat q_i - 0\right| = 0 \\
(\forall i > m) \qquad \left| \hat q_i - \frac{1}{m}\right| \leq \frac{1}{2m} \\ %\; \text{, taking the union bound and assuming } h \geq 24m \log\frac{4N}{\delta}\\
\end{cases}\\
\implies &
\begin{cases}
(\forall i \leq m) \qquad  \hat q_i  = 0 \\
(\forall i > m) \qquad \hat q_i \in [\frac{1}{2m}, \frac{3}{2m}]\\
\end{cases}\\
\implies & \hat{m} \leq 2m
}

Given $\boldsymbol{q} = \boldsymbol{q}^{b}$, we have by Lemma \ref{lem:conc1}, with probability at least $1 - \delta$ that:

\eq{
& \left| \hat q_i - q_i\right| 
 \leq \sqrt{\frac{6}{mh} \log\frac{2}{\delta}} \qquad \text{if } i \leq m \\
\implies &
(\forall i \leq m) \qquad  \hat q_i \leq \frac{3}{2m} \\
\implies & \hat{m} \geq \frac{2m}{3}
}

\end{proof}

\begin{proof}[Proof of \cref{thm:simple-regret}]

for $(i,j)\in A$, the algorithm explicitly selects the action, $X_i = j$,  $\frac{h}{2\hat m}$ times.
\eq {
\hat \mu_{i,j} = \frac{2\hat m}{h} \sum_{t=1}^{h/2\hat m} r_t(X_i = j)
}

Via Hoeffding's Inequality

\eq {
\P{ \left|\hat \mu_{i,j} - \mu_{i,j}\right| > \epsilon } \leq 2\exp{-\frac{h\epsilon^2}{\hat{m}}}
}

for $(i,j)\notin A$, the algorithm has observed the reward given $X_i = j$ at least $\frac{h}{2\hat m}$ times.

\eq {
(i,j)\notin A & \implies \hat{s}_i \geq \frac{1}{\hat m} \\
& \implies \sum_{t=1}^h \ind{X_i = j} \geq \frac{h}{\hat{m}}
}

Let  $Z_{ij} = \sum_{t=1}^{h/2} \ind{X_i = j}$ and $t'_1 ... t'_{Z_{ij}} = {t:X_{i,t} = j}$

\eq {
\hat \mu_{i,j} = \frac{1}{Z_{ij}}\sum_{t'=1}^{Z_{ij}}r_{t'}
}

\eq {
\P{\left| \hat \mu_{i,j} - \mu_{ij} \right| > \epsilon} = & \sum_{z=1}^\infty \P{Z_{ij}=z}\P{\left| \frac{1}{Z_{ij}} \sum_{t'=1}^{Z_{ij}}r_{t'} - \mu_{ij} \right| > \epsilon | Z_{ij} = z} \\
= & \sum_{z=1}^\infty \P{Z_{ij}=z}\P{\left| \frac{1}{z} \sum_{t'=1}^{z}r_{t'} - \mu_{ij} \right| > \epsilon } \\
\leq & \P{\left| \frac{2\hat m}{h} \sum_{t'=1}^{h/2\hat m}r_{t'} - \mu_{ij} \right| > \epsilon } \sum_{z=1}^\infty \P{Z_{ij}=z} \\
\leq & 2\exp{-\frac{h\epsilon^2}{\hat{m}}}\\
}

Applying the union bound over all $2N$ actions.
\eq{
& \P{\exists (i,j): \left| \hat \mu_{i,j} - \mu_{i,j} \right| > \epsilon} \leq 4N\exp{-\frac{h\epsilon^2}{\hat{m}}} \\
\implies & \P{\exists (i,j): \left| \hat \mu_{i,j} - \mu_{i,j} \right| > \sqrt{\frac{\hat m}{h}\log{\frac{4N}{\delta}}}} \leq \delta \\
}

Now by \cref{lem:m_est}, 
\eq{
h \geq 24m \log \frac{4N}{\delta} \implies \P{\hat m > 2m} \leq \delta
}

Therefore if $h \geq 24m \log \frac{4N}{\delta}$, we have with probability at least $1 - 2\delta$ that

\eqn{
\label{eqn:joint_bound_estimates_m}
(\forall i, j) \qquad \left|\hat \mu_{i,j} - \mu_{i,j}\right| \leq \sqrt{\frac{2m}{h} \log\frac{4N}{\delta}}\, \text { and } \hat{m} \leq 2m
}
Suppose $h < 24m \log \frac{4N}{\delta}$. Then
\eq{
\left|\hat \mu_{i,j} - \mu_{i,j}\right| \leq 1 \leq \sqrt{\frac{2m}{h} \log \frac{4N}{\delta}}\,.
}
Therefore
\eq{
\left|\hat \mu_{i,j} - \mu_{i,j}\right| \leq \sqrt{\frac{24m}{h} \log \frac{4N}{\delta}} \qquad \forall h\ \text{ with probability at least } 1-2\delta
}
Therefore 
\eq{
R_s(h) 
&\leq 2\delta + \sqrt{\frac{24m}{h} \log \frac{4N}{\delta}} \\
&\leq \frac{8m}{h} + \sqrt{\frac{24m}{h} \log \left(\frac{Nh}{m}\right)}
}
as required.
\end{proof}
